Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-01-20 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@

p <- ggplot2::ggplot(
  data = data,
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method))
p <- p + ggplot2::theme_light(base_size=base_size)
p <- p + ggplot2::geom_point()
p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
p <- p + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=1.5,
  hjust=1.0)
p <- p + ggplot2::scale_x_continuous(
  name = latex2exp::TeX("$\\mu$"),
  labels = scales::math_format())

p <- p + ggplot2::scale_y_continuous(
  name = latex2exp::TeX("$\\lambda$"))

p
```

@@

@@

p <- ggplot2::ggplot(
  data = data,
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method,
    shape = version))
p <- p + ggplot2::theme_light(base_size=base_size)
p <- p + ggplot2::geom_point()
p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
p <- p + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=1.5,
  hjust=1.0)
p <- p + ggplot2::scale_x_continuous(
  name = latex2exp::TeX("$\\mu$"),
  labels = scales::math_format())

p <- p + ggplot2::scale_y_continuous(
  name = latex2exp::TeX("$\\lambda$"))

p
```

>>>>

# Set general theme.
plot_theme <- ggplot2::theme_light(base_size=base_size)

@@

p_bc <- ggplot2::ggplot(
  data = data[method == "Box-Cox"],
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method))
p_bc <- p_bc + plot_theme
p_bc <- p_bc + ggplot2::geom_point()
p_bc <- p_bc + ggplot2::geom_hline(
  yintercept = 1.0, 
  linetype="longdash", 
  colour="grey40")
p_bc <- p_bc + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=-1.0,
  hjust=1.0)
p_bc <- p_bc + ggplot2::scale_x_continuous(
  name = latex2exp::TeX("$\\mu$"),
  labels = scales::math_format())
p_bc <- p_bc + ggplot2::scale_y_continuous(
  name = latex2exp::TeX("$\\lambda$"),
  limits = c(-3.6, 1.1))
p_bc <- p_bc + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")

p_yj <- ggplot2::ggplot(
  data = data[method == "Yeo-Johnson"],
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method))
p_yj <- p_yj + plot_theme
p_yj <- p_yj + ggplot2::geom_point()
p_yj <- p_yj + ggplot2::geom_hline(
  yintercept = 1.0,
  linetype="longdash",
  colour="grey40")
p_yj <- p_yj + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=-1.0,
  hjust=1.0)
p_yj <- p_yj + ggplot2::scale_x_continuous(
  name = latex2exp::TeX("$\\mu$"),
  labels = scales::math_format())
p_yj <- p_yj + ggplot2::scale_y_continuous(
  name = latex2exp::TeX("$\\lambda$"),
  limits = c(-3.6, 1.1))
p_yj <- p_yj + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE)
p_yj <- p_yj + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

p <- egg::ggarrange(plots=list(p_bc, p_yj), ncol=2, draw=FALSE)
p
```

@@

.create_lambda_shift_data <- function(x, shift_range){

@@

  return(data)
}

.create_density_plot <- function(x, plot_theme){
  p <- ggplot2::ggplot(
    data = data.table::data.table("x"=x),
    mapping = ggplot2::aes(
      x = x))
  p <- p + plot_theme
  p <- p + ggplot2::geom_density()
  p <- p + ggplot2::theme(
    axis.title = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank())

  return(p)
}

.create_lambda_shift_plot <- function(data, plot_theme, limits, expected=FALSE){

  annotation_settings <- get_annotation_settings(plot_theme)

  p <- ggplot2::ggplot(
    data = data,
    mapping = ggplot2::aes(
      x = mu,
      y = lambda,
      colour = method,
      shape = version))
  p <- p + plot_theme
  p <- p + ggplot2::geom_point()
  p <- p + ggplot2::geom_hline(
    yintercept = 1.0, 
    linetype="longdash", 
    colour="grey40")
  p <- p + ggplot2::scale_x_continuous(
    name = latex2exp::TeX("$\\mu$"),
    labels = scales::math_format())
  p <- p + ggplot2::scale_y_continuous(
    name = latex2exp::TeX("$\\lambda$"),
    limits = limits)

  if(expected){
      p <- p + ggplot2::annotate(
    geom = "text",
    x = 6,
    y = 1,
    label = "expected",
    colour=annotation_settings$colour,
    family=annotation_settings$family,
    fontface=annotation_settings$face,
    size=annotation_settings$geom_text_size,
    vjust=-1.0,
    hjust=1.0)
  }

  return(p)
}

# Normal distribution.
x <- stats::rnorm(10000L)
shift_range <- 10^seq(from=0, to=6, by=0.1)

# Right skewed data
x_right_skewed <- (x - min(x) + 1)^0.1
x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)

# Left skewed data
x_left_skewed <- (x - min(x))^3
x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)

#### Normal distribution -------------------------------------------------------

data <- .create_lambda_shift_data(
  x = x,
  shift_range = shift_range)

# Density
p_dens_normal <- .create_density_plot(
  x=x,
  plot_theme=plot_theme)

# Box-Cox
p_bc_normal <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(-3.6, 1.1))
p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_normal <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(-3.6, 1.1))
p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_yj_normal <- p_yj_normal + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

#### Right skewed distribution -------------------------------------------------

data <- .create_lambda_shift_data(
  x = x_right_skewed,
  shift_range = shift_range)

# Density
p_dens_right <- .create_density_plot(
  x = x_right_skewed,
  plot_theme = plot_theme)

# Box-Cox
p_bc_right <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(1.0, 65))
p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_right <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(1.0, 65))
p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE)
p_yj_right <- p_yj_right + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

#### Left skewed distribution --------------------------------------------------

data <- .create_lambda_shift_data(
  x = x_left_skewed,
  shift_range = shift_range)

# Density
p_dens_left <- .create_density_plot(
  x = x_left_skewed,
  plot_theme = plot_theme)

# Box-Cox
p_bc_left <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(-10.5, 0.5))
p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_left <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(-10.5, 0.5))
p_yj_left <- p_yj_left + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_yj_left <- p_yj_left + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

p <- egg::ggarrange(
  plots=list(
    p_dens_normal, p_bc_normal, p_yj_normal,
    p_dens_right, p_bc_right, p_yj_right,
    p_dens_left, p_bc_left, p_yj_left),
  ncol=3,
  draw=FALSE)
p
```
